Improving Item-Item Similarity Estimation during Collaborative Filtering

### 2014-03-25 paper/paper.tex
Touched up intro and score-based section

In our first approach, we model the probability distributions of true and 
observed similarity. The goal is to use the observed similarity score directly 
to approximate the true similarity. In this way, we can consider an abstact 
$TrueSim$ which ignores the details of $TrueSimRand$ or $TrueSimPref$. 
Similarly, we consider $Sim_n$ to be a generic observed similarity score when
the two itmes have $n$ common reviewers. Now, let $Y$ be a random variable 
representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
random variable representing the $Sim_{n}$ of a pair of items when $n$ common 
users are observed.

@@

The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
and $Y$ ($TrueSim$). Moreover, the best predicted $TrueSim$ is a linear 
combination of the average true similarity and the observed similiarity. Since $\sigma_{err,n}^2$ is 
nonnegative, the slope is always less than or equal to one, meaning that the 
observed distance from $\mu$ (the average true similarity) overestimates the 
true distance. This is reasonable, since prior suggests that the true 
similarity is likely to be near $\mu$.\footnote{It is worth noting that despite
the resemblance to Slope One Predictors \cite{Lemire1998}, this is a fundamentally 
different technique. Here, we have a linear relationship between observed 
scores and true scores, whereas the Slope One model seeks a linear relationship 
between the ratings a user gives to one items versus another.}

This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.
To gain a more generality, we model $\sigma_{err,n}^2$ as a function of $n$
instead of considering them independent of each other. Recall that 
$\sigma_{err,n}^2$ is the noisiness of the observed $Sim_n$ about $TrueSim$. 
Since the $TrueSim$ is the $Sim$ when all users that will rate the items 
have rated them, $Sim_n$ can be thought of as sampling $n$ users from the 
set used to calculate $TrueSim$. Although $Sim$ is likely more complicated 
than an average, the Central Limit Theorem motivates the intuition that the 
variance of $Sim$ would decrease as the sqaure root of $n$. Thus we model 
$\sigma_{err,n}^2$ as:
\begin{align}
\sigma_{err,n}^2 = \frac{\alpha}{\sqrt{n}},
\end{align}
yielding the single parameter model:
\begin{align}
\left(Y - \mu\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
\left(X_n-\mu\right)
\end{align}

@@

With a measurable $GoldSim$ as our goal, we can apply supervised learning 
techniques. First, we find pairs of items that have at least $N$ common users 
and calculate their $GoldSim$. Next we withold some of the users to produce 
estimated $Sim_n$ for $n=1,...,N$. By doing this over many pairs of items, we 
produce pairs of ($Sim_n$, $TrueSim$) that can be used for training and testing.

Now, the parameters $\mu$, $\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be 
approximated using the method of moments over the training data. The values of 
$\sigma_{err,n}^2$ are fit to the function $\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ 
that minimizes the sum of squared error is computed.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear models. For each number of reviewers
    in common $n$, the graphs represent the average squared error of predicted
    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
    of guessing the average $GoldSim$ is included.}
    \label{fig:Linear}
\end{figure}

>>>>

In our first approach, we model the probability distributions of true and 
observed similarity. The goal is to use the observed similarity score directly 
to approximate the true similarity. In this way, we can consider an abstact 
$TrueSim$ which ignores the details of $TrueSimRand$ or $TrueSimPref$. 
Similarly, we consider $Sim_n$ to be a generic observed similarity score when
the two items have $n$ common reviewers. Now, let $Y$ be a random variable 
representing the $TrueSim$ of a randomly chosen pair of items. Let $X_{n}$ be a
random variable representing the $Sim_{n}$ of a pair of items when $n$ common 
users are observed.

@@

The model suggests that there is a linear correlation between $X_n$ ($Sim_n$)
and $Y$ ($TrueSim$).\footnote{It is worth noting that despite
the resemblance to Slope One Predictors \cite{Lemire1998}, this is a fundamentally 
different technique. Here, we have a linear relationship between observed 
scores and true scores, whereas the Slope One model seeks a linear relationship 
between the ratings a user gives to one items versus another.} 
Moreover, the best predicted $TrueSim$ is a linear 
combination of the average true similarity and the observed similiarity. Since $\sigma_{err,n}^2$ is 
nonnegative, the slope is always less than or equal to one, meaning that the 
observed distance from $\mu$ (the average true similarity) overestimates the 
true distance. This is reasonable, since the prior suggests that the true 
similarity is likely to be near $\mu$.

This model requires a different parameter $\sigma_{err,n}^2$ for each $n$.
To gain more generality, we model $\sigma_{err,n}^2$ as a function of $n$
instead of considering them independent of each other. Recall that 
$\sigma_{err,n}^2$ is the noisiness of the observed $Sim_n$ about $TrueSim$. 
Since $TrueSim$ is $Sim$ when all users that will rate the items 
have rated them, $Sim_n$ can be thought of as sampling $n$ users from the 
set used to calculate $TrueSim$. Although $Sim$ is usually more complicated 
than an average, the Central Limit Theorem motivates the intuition that the 
variance of $Sim$ would decrease as the sqaure root of $n$. Thus we model 
$\sigma_{err,n}^2$ as:
\begin{align}
\sigma_{err,n}^2 = \frac{\alpha}{\sqrt{n}},
\end{align}
yielding the single parameter model:
\begin{align}
\left(Y - \mu\right) = \frac{\sigma_{s}^2}{\sigma_{s}^2+\frac{\alpha}{\sqrt{n}}}
\left(X_n-\mu\right)
\end{align}

@@

With a measurable $GoldSim$ as our predictive goal, we can apply supervised 
learning techniques. First, we find pairs of items that have at least $N$ common
users and calculate their $GoldSim$. Next we withold some of the users to 
produce estimated $Sim_n$ for $n=1,...,N$. By doing this over many pairs of 
items, we produce pairs of ($Sim_n$, $TrueSim$) across different values of $n$ 
that can be used for training and testing. Now the parameters $\mu$, 
$\sigma_{s}^2$, and $\sigma_{err,n}^2$ can be approximated using the method of 
moments over the training data. The values of $\sigma_{err,n}^2$ are fit to the 
function $\frac{\alpha}{\sqrt{n}}$ and the $\alpha$ that minimizes the sum of 
squared error is computed.

@@

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.7\textwidth]{LinearTechniques.png}
	\caption{Comparison of different linear techniques. For each number of reviewers
    in common, $n$, the graphs represent the average squared error of predicted
    $GoldSim$ calculated from the observed $Sim_n$. As a baseline, the technique
    of guessing the average $GoldSim$ is included. For these experiments,
    $N=100$ for determining $GoldSim$ and $70\%$ of the data was used for
    training.}
    \label{fig:Linear}
\end{figure}
